Influence of journal data archiving mandates on author attitudes, experiences, and practices

### 2018-11-22 manuscript.Rmd
add a draft of the abstract

---
title: Influence of journal data archiving mandates on author attitudes, experiences,
  and practices
author: "Heather A. Piwowar, Todd J. Vision"
csl: peerj.csl
output:
  word_document:
    fig_caption: yes
  md_document:
    variant: markdown_github
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  html_document:
    css: style.css
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
bibliography: input_references.bib
---

@@

The invitations were sent out monthly between February 2011 and March 2014 to a total of 23853 23853  authors, and 4102 4102 responded (corresponding to a response rate of 17%).

the introduction of the JDAP data archiving policy significantly increases the odds of a positive response regarding data sharing by a factor of 1.36
1.36
 (CI=[1.16;1.60]
1.16
1.60
), while publishing date and whether the journal ever adopted a JDAP policy did not show any statistically significant effects.

We found the similar results when we looked at the author’s responses to community norms about public data archiving rather than peer-to-peer data sharing

this study only facilitates a relatively short-term look at policy effect: it will likely take many more years for the effect of these policies to be fully felt.

>>>>

---
title: Influence of journal data archiving mandates on author attitudes, experiences, and practices
author: "Heather A. Piwowar, Todd J. Vision"
csl: peerj.csl
output:
  html_document:
    css: style.css
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
  md_document:
    variant: markdown_github
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
bibliography: input_references.bib
---

@@

INTRODUCTION: Some journals require that their authors share the raw data behind their analyses, but little is known about how the introduction of these mandates affects the attitudes, experiences, and behaviour of authors.

OBJECTIVE: This study measures the correlation of journal data sharing policies with the data sharing attitudes, experiences, and behaviour of authors.

METHODS: An online questionaire was sent to all corresponding authors of 48 journals in the fields of evolution, ecology, and general biology who published papers between November 2010 and March 2014. The online questionnaire was designed to measure the respondent’s practices, experiences, and opinions around public archiving of research datasets by asking about their opinion of community norms, recent behaviour with respect to a recently published paper, previous experiences with data sharing, and attitudes toward hypothetical journal policies.

RESULTS:  Survey invitations were sent out to 23,853 corresponding authors; 4,102 responded (response rate of 17%).  Twenty-one of the journals had adopted a journal sharing requirement by the time the survey was completed.  We found the introduction of a data archiving policy significantly increased the respodent's belief that data sharing was a community norm, whereas publishing date and whether the journal ever adopted a data sharing policy did not show any statistically significant effects.  In exploratory analysis, the introduction of a journal data policy was correlated with a decrease in worries about "scooping", and increased the feeling that data archiving policies was "the right thing to do."

CONCLUSION: This study suggests that although journals may be afraid that data archiving policies will be met with resentment and a backlash from their authors, the attitudes of autohrs after policies go into effect is positive and the policies serve to increase and normalize data sharing within their disciplines.

DATA AVAILABILITY:  The data from this study is publicly available.
